---
title: "PCA for Fluid analysis data"
output:
  html_document:
    df_print: paged
---

Based on Paper that published: https://iopscience.iop.org/article/10.1088/1757-899X/598/1/012088/meta

# Load Library

```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(openxlsx)
library(factoextra)
library(hrbrthemes)
library(ggfortify)
library(corrplot)
library(ggcorrplot)
library(ggbiplot)
library(RColorBrewer)
```

# Open & wrangling dataset

Dataset are from fluid sample analysis of Heavy dump truck 100-mton class for Transmission component data.

```{r}
read.xlsx("raw_data/777_Trans.xlsx", sheet = "Trans") %>%
  glimpse() -> raw_data

raw_data %>%
  select(
    unitno,
    Si:B,
    V100, PQ,
    PC06:ISO.14,
    oilchanged2:filterchanged2,
    Filter.Cut:Magnetic,
    life_hrs:RUL
  ) -> df_trans

glimpse(df_trans)
```
After filtering for Variables that needed for PCA Processing and save in object name as *df_trans* then we calculate the missing data in the dataset

```{r missing data}
df_trans %>%
  summarise_all(~sum(is.na(.))) %>%
  glimpse()

```

So, we select the completed data from the missing data
```{r completed cases}
df_trans %>%
  drop_na() %>%
  mutate(
    oilchanged2 = factor(oilchanged2, levels = c(1,0)),
    filterchanged2 = factor(filterchanged2, levels = c(1,0)),
    Filter.Cut = factor(Filter.Cut, levels = c("A","B","C","X"), labels = c(1,2,3,4)),
    Magnetic = factor(Magnetic, levels = c("A","B","C","X"), labels = c(1,2,3,4))
  ) %>% glimpse() -> df_fix

df_fix


```

Oil Changed, Filter Changed, Filter Cut and Magnetic are Factorial variable and wee need modify the structure. and then we checked for the correlation

```{r correlation matrix}

df_fix %>%
  select(-c(unitno)) %>%
  mutate(
    oilchanged2 = as.numeric(oilchanged2),
    filterchanged2 = as.numeric(filterchanged2),
    Filter.Cut = as.numeric(Filter.Cut),
    Magnetic = as.numeric(Magnetic),
  ) %>%
  cor() %>%
  corrplot(method = "color",
           type = "lower",
           tl.col = "black", tl.srt = 1,
           addCoef.col = "black",
           number.cex = 0.55,
           col=brewer.pal(n = 10, name="RdYlBu"),
           mar=c(.25,0,.25,.25))

```



when we drop the missing data, we can normalized the data before processing into PCA. Normalization or Scaling data in R can be using with *scale* function that built in R which is Z-score normalization.

```{r z-score}

  df_fix %>%
  select(-c(unitno, oilchanged2:RUL)) %>%
  as_tibble() %>%
  scale(center = TRUE, scale = TRUE)
  
```
```{r minmax}
norm_minmax <- function(x){
                           (x- min(x)) /(max(x)-min(x))
}

df_fix %>%
  select(-c(unitno, oilchanged2:RUL)) %>%
  as_tibble() %>%
  norm_minmax()
```

PCA analysis with *prcomp()* package in R

```{r prcomp}

df_fix %>%
  select(-c(unitno, life_hrs:RUL)) %>%
  mutate(
    oilchanged2 = as.numeric(oilchanged2),
    filterchanged2 = as.numeric(filterchanged2),
    Filter.Cut = as.numeric(Filter.Cut),
    Magnetic = as.numeric(Magnetic),
  ) %>%
  prcomp(scale. = TRUE, center = TRUE) -> pca_trans

pca_trans

```
Create a screeplot to find how much *n*

```{r screeplot}
pca_trans %>%
  screeplot(type = "line", npcs = 20, ylim = c(0,5)) +
  abline(h = 1, lty = 3, col = "red") # based on scree plot pca n = 6
```
```{r summary PCA}
summary(pca_trans)
```

Plotting of Individual PC is representative percentage for variables in dataset.

```{r}
pca_trans %>%
  fviz_eig(addlabels = TRUE, ncp = 20,
           barfill = "springgreen3",
           ggtheme = theme_ipsum_rc(
             ticks = TRUE,
             grid = FALSE
           ))
```

```{r biplot}
pca_trans %>%
  biplot(scale = 0, cex = .75)

```

Another thing you can do to easily see the grouping of different variables is a biplot. With this plot could be found that It’s important to know what your components are made up of. In other words, which variables contribute to any given component.


```{r biplot advance}
pca_trans %>%
  autoplot(data = df_fix, colour = 'life_hrs',
           loadings = TRUE, loadings.label = TRUE, loadings.colour = "purple", loadings.label.size = 3, scale = 1,
           loadings.label.vjust = 1.5) +
  scale_color_gradient(low="springgreen", high="firebrick2") +
  theme_ipsum_rc(
    base_size = 10,
    plot_title_size = 12,
    subtitle_size = 10,
    strip_text_face = "bold",
    grid = TRUE,
    ticks = FALSE
  )
```

```{r}
print(pca_trans$rotation)
```
If we want to analysis top-12 PC in dataset.

```{r}
print(pca_trans$rotation[,1:12])
```


```{r}
pca_trans %>%
  ggbiplot(
    obs.scale = 1,
    var.scale = 1,
    groups = df_fix$life_hrs,
    ellipse = TRUE,
    circle = TRUE,
    ellipse.prob = 0.68
  ) + 
  # scale_color_discrete(name = '') +
  scale_color_gradient(low="springgreen", high="firebrick2", name = "Life Hours") +
  theme_ipsum_rc(
    base_size = 10,
    plot_title_size = 12,
    subtitle_size = 10,
    strip_text_face = "bold",
    grid = FALSE,
    ticks = TRUE
  ) +
  theme(
    legend.direction = 'horizontal',
    legend.position = 'top'
  )
```

